GVN-Hoist: Hoisting Computations from Branches

### 2016-08-08 hoist.tex
add more text on introduction

Classical scalar optimizations like CSE \cite{dragonbook} work very well on
single basic blocks.  With time, we have seen the transition to more complex
passes like GCSE that work across basic blocks, or other passes like the
macroblock formation that create a region within

Why hoisting expressions:
\begin{itemize}
\item OOO processors: good for speculative execution of branches
\item more ILP for in-order processors
\item code size
\item optimize for register pressure
\item improve inlining heuristics
\end{itemize}

>>>>

Compiler techniques to remove redundant computations are composed of an analysis
phase that detects identical computations in the program and a transformation
phase that reduces the number of run-time computations.  Classical scalar
optimizations like CSE \cite{dragonbook} work very well on single basic blocks.
When it comes to detect redundancies across basic blocks these techniques fall
short: more complex passes like GCSE and PRE have been designed to handle these
cases based on dataflow analysis \cite{morel1979global}.  At first these
techniques were described in the classical data-flow analysis framework, and
later the use of the SSA representation lowered the cost in terms of compilation
time \cite{briggs1994effective,chow1997new,kennedy1999partial} and brought these
techniques in the main stream: nowadays SSA based PRE is available in every
industrial compiler.

This paper describes code hoisting, a technique that uses the information
computed for PRE to detect identical computations and has a transformation phase
whose goal differs from PRE: it removes identical computations from different
branches of execution.  These identical computations would not be redundant at
run-time and thus the number of run-time computations is not reduced.  Instead,
the goals of code hoisting are:
\begin{itemize}
\item to reduce the code size of the program,
\item to expose more instruction level parallelism by hoisting identical
  computations to be executed earlier, and thus helping to schedule heavy
  computations earlier in order to avoid pipeline bubbles,
\item to help out-of-order processors with speculative execution of branches: by
  hoisting expressions out of branches, code hoisting can effectively reduce
  the amount of code to be speculatively executed,
\item to decrease the number of live variables and thus to reduce register
  pressure,
\item to improve function inlining heuristics by reducing the size of a function body.
\end{itemize}
